### 2020-04-20 README.Rmd
code review setup

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# berlin.trees

 <img src="analysis/figures/map_03_uhi.png" align="right" height = 275px, width = 275px />

## Purpose

This repository contains an analysis of tree growth (potential) and its relationship with the Urban Heat Island (UHI) effect in Berlin using an extensive, publicly available data set. 
For some details of the (on-going) research, download this [file](./analysis/paper/paper.html) and have a wee read if you like.
Having a look at this first (mainly figures) will help contextualize some of the workflow/code/results, I think!

## Approach

The folder structure follows the principal of a [**research compendium**](https://research-compendium.science/) following concepts developed by [Ben Marwick and others](https://www.tandfonline.com/doi/abs/10.1080/00031305.2017.1375986?journalCode=utas20).
That means all analyses and the final write-up are organized in a (fairly) coherent manner, encouraing reproducibility.

I chose to package up the analyses in an `R` package to facilitate reproducibility, as dependencies are more easily tracked, and analyses steps can be broken down into functions that can be tested, adjusted when necessary, and most important, re-used.
A brief overview and benefits of that approach are detailed in my [blog post here](https://the-hull.github.io/reproducibility_compendium/analyses/00_write-up.html).

A workflow-management software [**`drake`**](https://github.com/ropensci/drake) and approach is also something I've relied on heavily here, as the analysis is quickly growing.
While it takes time to get used to, and to implement, it has already paid off massively when I had to go back and change parameters/inputs.
`drake` is extensively documented ([see book](https://books.ropensci.org/drake/)), and I highly recommend it.

In brief:

- Research compendium for coherent structure (based on `R` package; recognizable by others)
- Packaged analyses steps/functions into `R package` (`berlin.trees`; call as `berlin.trees::my_function()` in scripts)
- employed `drake` for managing analyses workflow

Below is a brief explanation of the repository structure, and important files, as well as what drake does and how it relates to the analysis.

### Repository structure

The repo structure follows that of a typical `R` package.
There are:

- high-level files (this readme, `NAMESPACE`, `DESCRIPTION`) relating to typical `R` package development
- high-level files relating to `drake` and its implementation (`make.R`, `_drake.R`)
- high-level files that I use to get going with a package/compendium (`.Rprofile`, `_setup.R`)

- `R` folder for functions
- `man` folder for the documentation (i.e. help files) of the functions that pop-up in the Help viewer in `RStudio` (think of future you!)
- `analysis` that containts figures and the manuscript as a Rmarkdown / latex document

- output papers from the "raw" (uncompiled) manuscript (`paper_knit.xyz`)

### Using `drake`

`drake` uses an analyses plan (a set of instructions) to generate results. 
The great thing about `drake` is that it knows which instructions depends on another (i.e. download data before plotting it, than put it in a manuscript).
The set of instructions are located in `R/plan.R`, and outline all the steps that I take to generate results, figures and rendering the final manuscript.

The files `make.R` and `_drake.R` both achieve the same thing, but in different contexts (typically difference is running the pipeline on a local `R` install vs. run on a server). 
If `plan.R` is the brain, the above files are the muscles making things move. 
Note, that they are generated automatically by `drake`, so you don't need to worry that much about what to put in them (`drake::use_drake()`).

The file `_setup.R` is typially what I work in to start the pipeline, or add libraries to my analysis, and it contains the function `drake::r_make()` that starts of the pipeline!

#### The `plan.R`

`plan.R` is structured following the entire analyses from downloading tabular and spatial data, cleaning it, producing stats and plots, and then compiling the manuscript.
I typically seperate indivudal steps with `# Step -----------` as that gives you a neat outline in `RStudio` (if click the lines symbol above the scroll-bar in the script window).
Every analyses step is named, and that name can then be used in a sub-sequent step (this is how drake manages dependencies).
For example, see this taken directly from the documentation for a fairly self-explanatory application:

![](https://camo.githubusercontent.com/e9cd15743788ea28d336848828f458aad86f1ad9/68747470733a2f2f646f63732e726f70656e7363692e6f72672f6472616b652f7265666572656e63652f666967757265732f67726170682e706e67)

`drake` caches (intermediate) results and keeps track of any changes to them, and tells you which parts steps will be re-executed next time to finish of the pipeline! 
Super cool. 
Any intermediate result can, for example, be inspected using `drake::readd(X)` or `drake::loadd()`, and this is also how you include results in e.g. an `Rmarkdown` report.

### Workflow

Note, that I've added some descriptions to the steps in `plan.R`, but typically I rely on the function help that I generate in `R/functions.R`. 
The workflow I apply there is typically iterative, so I play around in a few dump/script files (`analyses/scripts/...`) and then refactor the code for a specific use into a function.
As I've bundled the analysis into a package, any (properly formatted) documentation in the functions is then accessible using `berlin.trees::my_function()`, and should I forget what a function does, I put my cursor on the function name and hit `F1` - the help then appears in the `RStudio` viewer. 
I try to use verbs and wordy function names so I hope that most things would be self-explanatory (at least for myself).
This also works for documenting data sets (e.g. try `?iris` in your `R` console) - and can be super helpful!

Once functions are coded and the plan configured, I typically navigate to the file `_setup.R` (either using the shortcud `CTRL+.` or clicking) to run `drake::r_make()` and a few other things, or I just execute the latter command directly in the console. 
`drake` then let's you know what's up!

### Code

As mentioned before, I typically break down analyses into many steps, and use wordy names to indicate what they do. 
I haven't implemented formal unit testing here, but I'm quite thorough with checking outputs (especially in exploratory analyses), and doing sanity checks on outputs.
With the break-down approach, things tend to stay fairly straightforward for every step, and I only comment code where entirely necessary (for myself so far).
I do try to keep the functions organized (using `# Section ----------` - ), so that an outline appears in `RStudio`:

![](./functions.png)

That really helps navigating!

### Review / Feedback

As Tom said, don't spend too much time going through individual things!
I think having a look at the structure, and whether it's "straightforward" to follow along would be most useful, especially given that there are two potentially new concepts here (`R package` development and `drake`)
Thanks for having a go!!

<!-- [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio) -->

<!-- This repository contains the data and code for our paper: -->

<!-- > Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx> -->

<!-- Our pre-print is online here: -->

<!-- > Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

<!-- ### How to cite -->

<!-- Please cite this compendium as: -->

<!-- > Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

<!-- ### How to download or install -->

<!-- You can download the compendium as a zip from from this URL: </archive/master.zip> -->

<!-- Or you can install this compendium as an R package, berlin.trees, from GitHub with: -->

<!-- ```{r gh-installation, eval = FALSE} -->
<!-- # install.packages("devtools") -->
<!-- remotes::install_github("/") -->
<!-- ``` -->

<!-- ### Licenses -->

<!-- **Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/) -->

<!-- **Code :** See the [DESCRIPTION](DESCRIPTION) file -->

<!-- **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->

<!-- ### Contributions -->

<!-- We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms. -->