### 2022-10-11 analysis/paper/author_statement_credit.Rmd
reading submission

---
title: "CRediT author statement"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style

---

Following https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement

Alexander Hurley: Conceptualization, Methodology, Data curation, Formal analysis, Visualization, Writing - Original Draft.
Ingo Heinreich: Resources, Investigation, Project administration, Funding acquisition, Writing - Review & Editing